---
title: "p8105_hw2_yl4610"
author: "Yuxin Liu"
output: github_document
date: "2022-10-01"
---

#question1
```{r setup, message=FALSE, warning = FALSE}
# load the libraries
library(tidyverse)
library(readxl)
library(dplyr)
# Read and clean the data
NYCtransit_data = read_csv( "./datafiles/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>%
janitor::clean_names() %>% 
# retain certain variables
select(line, station_name, station_latitude, station_longitude, route1:route11, entry, exit_only, vending, entrance_type, ada) %>% 
# Convert the entry variable from character (YES vs NO) to a logical variable
mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```
This dataset contains variables such as `r names(NYCtransit_data)`. 
I used read_csv to read the dataset. Then I used janitor to clean names, select to include variables that I want, and mutate to convert the character variable "entry" to logical variable.  
It contains `r ncol(NYCtransit_data)` columns and `r nrow(NYCtransit_data)` rows.
These data are not tidy because route number should be a variable instead of having 11 columns of different routes. I should convert `route` variables from wide to long format. In this way, it is more convenient to check on specific routes within one column.

How many distinct stations are there?
```{r}
NYCtransit_data %>% 
  select(station_name, line) %>% 
  distinct
```
There are 465 distinct stations.

How many stations are ADA compliant?
```{r}
NYCtransit_data %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```
84 stations are ADA compliant.

What proportion of station entrances / exits without vending allow entrance?
```{r}
NYCtransit_data %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```
0.3770 or 37.70% station entrances / exits without vending allow entrance

```{r}
NYCtransit_data %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

NYCtransit_data %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```
There are 60 distinct stations serve the A train. Of the stations that serve the A train, 17 are ADA compliant.

#question2

Mr. Trash Wheel dataset:
```{r, message=FALSE, warning = FALSE}
# Read and clean the data
MrTrash_data = read_excel(
  "./datafiles/Trash_Wheel_Collection_Data.xlsx", 
# specify the sheet in the Excel file and to omit non-data entries
  sheet = 1,
  range = "A2:N549") %>%  
# use reasonable variable names
janitor::clean_names() %>%
# omit rows that do not include dumpster-specific data
drop_na(dumpster) %>% 
# round sports balls to nearest integer and converts it to an integer variable
mutate(
  sports_balls = round(sports_balls),
  sports_balls = as.integer(sports_balls),
# change year to numeric variables 
       year = as.numeric(year),
# add an additional variable
  from = "Mr"
  ) %>%
# reorganize 
relocate()
```

Professor Trash Wheel dataset:
```{r, message=FALSE, warning = FALSE}
# Read and clean the data
ProfTrash_data = read_excel(
  "./datafiles/Trash_Wheel_Collection_Data.xlsx", 
# specify the sheet in the Excel file and to omit non-data entries
  sheet = 2,
  range = "A2:M96") %>%  
# use reasonable variable names
janitor::clean_names() %>%
# omit rows that do not include dumpster-specific data
drop_na(dumpster) %>% 
# add an additional variable
mutate (from = "Professor") %>%
# reorganize 
relocate()
```
```{r}
# combine two datasets 
combine = bind_rows(MrTrash_data, ProfTrash_data)
```

For Mr. Trash Wheel dataset:
The names of variables are `r names(MrTrash_data)`. There are `r nrow(MrTrash_data)` rows and `r ncol(MrTrash_data)` columns.

For Professor Trash Wheel dataset:
The names of variables are `r names(ProfTrash_data)`. There are `r nrow(ProfTrash_data)` rows and `r ncol(ProfTrash_data)` columns.

```{r}
# filter to collect data in 2020
MrTrash_data_new = 
  MrTrash_data %>%
  filter (year==2020)
```

For combine dataset:
The names of variables are `r names(combine)`. There are `r nrow(combine)` rows and `r ncol(combine)` columns.Thus, there are 641 observations.
There are `r `observations in the resulting dataset. 
The total weight of trash collected by Professor Trash Wheel `r sum(pull(ProfTrash_data, weight_tons))`. The total number of sports balls collected by Mr. Trash Wheel in 2020 is `r sum(pull(MrTrash_data_new, sports_balls))`.

#question3

clean the data in pols-month.csv
```{r, message=FALSE, warning = FALSE}
#read pols-month.csv
pols_month_data = read_csv( "./datafiles/fivethirtyeight_datasets/pols-month.csv") %>%
janitor::clean_names() %>% 
# break up the variable mon into integer variables
separate(mon, into = c("year", "month", "day")) %>% 
#  replace month number with month name
mutate(month = month.name[as.numeric(month)],
 # change year to numeric variables 
       year = as.numeric(year),
# create a president variable taking values gop and dem
       president = ifelse(prez_dem == 1, "democratic", "republican")
       ) %>%
# remove prez_dem and prez_gop; and remove the day variable
select(-prez_dem, -prez_gop, -day)
```
clean the data in snp.csv 
```{r, message=FALSE, warning = FALSE}
#read snp.csv 
snp_data = read_csv( "./datafiles/fivethirtyeight_datasets/snp.csv") %>%
janitor::clean_names() %>% 
# reformat date 
mutate(date = as.Date(as.character(date), format = "%m/%d/%y")) %>%
# break up the variable mon into integer variables
separate(date, into = c("year", "month", "day")) %>%
#  replace month number with month name
mutate(month = month.name[as.numeric(month)],
# change year to numeric variables 
       year = as.numeric(year))%>%
# remove the day variable
select(-day) %>%
# organize so that year and month are the leading columns
relocate("year", "month")
```
tidy the unemployment data
```{r, message=FALSE, warning = FALSE}
#read unemployment.csv 
unemploy_data = read_csv( "./datafiles/fivethirtyeight_datasets/unemployment.csv") %>%
janitor::clean_names() %>% 
# switch from “wide” to “long” format
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment") %>% 
#  replace month number with month name
mutate(
  month = recode (month, "jan"=1, "feb"=2, "mar"=3, "apr"=4, "may"=5, "jun"=6, "jul"=7, "aug"=8, "sep"=9, "oct"=10, "nov"=11, "dec"=12),
  month = month.name[as.numeric(month)])

```
Join the datasets by merging snp into pols, and merging unemployment into the result
```{r}
threedatasets = 
 left_join(pols_month_data, snp_data, by = c("year", "month")) %>% 
 left_join(unemploy_data, by = c("year", "month"))
```

Names of variables in pols_month_data are `r names(pols_month_data)`. There are `r nrow(pols_month_data)` rows and `r ncol(pols_month_data)` columns. Dimension is `r dim(pols_month_data)` and range of years is `r range(pull(pols_month_data, year))`.

Names of variables in snp_data are `r names(snp_data)`. There are `r nrow(snp_data)` rows and `r ncol(snp_data)` columns.Dimension is `r dim(snp_data)` and range of years is `r range(pull(snp_data, year))`.

Names of variables in unemploy_data are `r names(unemploy_data)`. There are `r nrow(unemploy_data)` rows and `r ncol(unemploy_data)` columns.Dimension is `r dim(unemploy_data)` and range of years is `r range(pull(unemploy_data, year))`.

Names of variables in threedatasets are `r names(threedatasets)`. There are `r nrow(threedatasets)` rows and `r ncol(threedatasets)` columns.Dimension is `r dim(threedatasets)` and range of years is `r range(pull(threedatasets, year))`.