---
title: "p8105_hw2_yl4610"
author: "Yuxin Liu"
output: github_document
date: "2022-10-01"
---

#question1
```{r setup}
# load the libraries
library(tidyverse)
library(readxl)
library(dplyr)
# Read and clean the data
NYCtransit_data = read_csv( "./datafiles/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>%
janitor::clean_names() %>% 
# retain certain variables
select(line, station_name, station_latitude, station_longitude, route1:route11, entry, exit_only, vending, entrance_type, ada) %>% 
# Convert the entry variable from character (YES vs NO) to a logical variable
mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```
This dataset contains variables such as `r names(NYCtransit_data)`. 
I used janitor to clean names, select to include variables that I want, and mutate to convert the character variable "entry" to logical variable.  
It contains `r ncol(NYCtransit_data)` columns and `r nrow(NYCtransit_data)` rows.
These data are not tidy because route number should be a variable instead of having 11 columns of different routes. I should convert `route` variables from wide to long format. In this way, it is more convenient to check on specific routes within one column.

```{r}
NYCtransit_data %>% 
  select(station_name, line) %>% 
  distinct
```
There are 465 distinct stations.

How many stations are ADA compliant?
```{r}
NYCtransit_data %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```
84 stations are ADA compliant.

```{r}
NYCtransit_data %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```
0.3770 or 37.70% station entrances / exits without vending allow entrance

```{r}
NYCtransit_data %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

NYCtransit_data %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```
There are 60 distinct stations serve the A train. Of the stations that serve the A train, 17 are ADA compliant.

#question2
Mr. Trash Wheel dataset:
```{r}
# Read and clean the data
MrTrash_data = read_excel(
  "./datafiles/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
# specify the sheet in the Excel file and to omit non-data entries
  sheet = 1,
  range = "A2:N534") %>%  
# use reasonable variable names
janitor::clean_names() %>%
# omit rows that do not include dumpster-specific data
drop_na(dumpster) %>% 
# round sports balls to nearest integer and converts it to an integer variable
mutate(
  sports_balls = as.integer(sports_balls),
  dumpster = as.numeric(dumpster),
# add an additional variable
  Mr_Trash = "a"
  ) %>%
# reorganize 
relocate()
```

 Professor Trash Wheel dataset:
```{r}
# Read and clean the data
ProfTrash_data = read_excel(
  "./datafiles/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
# specify the sheet in the Excel file and to omit non-data entries
  sheet = 2,
  range = "A2:N116") %>%  
# use reasonable variable names
janitor::clean_names() %>%
# omit rows that do not include dumpster-specific data
drop_na(dumpster) %>% 
# round sports balls to nearest integer and converts it to an integer variable
mutate(
  sports_balls = as.integer(sports_balls),
# add an additional variable
  Prof_Trash = "b"
  ) %>%
# reorganize 
relocate()
```
```{r}
# combine two datasets 
inner_join(MrTrash_data, ProfTrash_data, by = "dumpster" )
```

The total weight of trash collected by Professor Trash Wheel? The total number of sports balls collected by Mr. Trash Wheel in 2020?


#question3
clean the data in pols-month.csv
```{r}
#read pols-month.csv
pols_month_data = read_csv( "./datafiles/fivethirtyeight_datasets/pols-month.csv") %>%
janitor::clean_names() %>% 
# break up the variable mon into integer variables
separate(mon, into = c("year", "month", "day")) %>% 
#  replace month number with month name
mutate(month = month.name[as.numeric(month)]) %>%
# create a president variable taking values gop and dem
mutate(president = prez_gop + prez_dem) %>%
# remove prez_dem and prez_gop; and remove the day variable
select(-prez_dem, -prez_gop, -day)
```
clean the data in snp.csv 
```{r}
#read snp.csv 
pols_month_data = read_csv( "./datafiles/fivethirtyeight_datasets/snp.csv") %>%
janitor::clean_names() %>% 
separate(date, into = c("month", "day", "year"))


```
# reformat date 
mutate(date = as.Date(as.character(date), format = "%d/%m/%y"))
# break up the variable date into integer variables
separate(date, into = c("year", "month", "day")) %>% 
#  reformat date 
mutate(date = month.name[as.numeric(month)]) %>%

# remove the day variable
select(-day) %>%
# organize so that year and month are the leading columns
relocate("year", "month")
